### 2024-01-25 parts/correction.tex
writing to publish

The counter example provided by \cref{def:bad-polynomial} relies on the fact
that $\CoveredPoly$ is not stable under fixing indeterminates, while
\kl{$\Nat$-polyregular functions} are. In this section, we prove that closing
$\CorrectPoly$ under assignments of variable is enough to recover from
\cref{karh:thm}. Because the previous attempt to characterize this class of
polynomials did not hold, this part of the paper will try to be extremely
precise.

@@

Let us now prove that \kl{$\Nat$-rational polynomials} are in $\CorrectPoly$.
The following \cref{n-poly-combinatorics:lem} is a slight refinement over the
classical combinatorial analysis of \kl{$\Rel$-polyregular functions}
\cite[Lemma 4.16]{LOPEZ23b} and \kl{$\Nat$-polyregular functions} that takes
into account \emph{all} \kl{maximal monomials} instead of a specific one
\cite[Lemma 5.37]{gaetanphd}.

\begin{lemma}
    \label{n-poly-combinatorics:lem}
    Let $f$ be a  \kl{$\Nat$-polyregular} function. 
    There exists $\omega \in \Nat$
    such that for all $p \in \Nat$,
    for all $\alpha_0, \dots, \alpha_p \in \Sigma^*$,
    for all $u_1, \dots, u_p \in \Sigma^*$,
    there exists a polynomial $P \in \Rel[X_1, \dots, X_p]$
    whose \kl{maximal monomials} are \kl{non-negative},
    and such that for all $X_1, \dots, X_p \geq \omega$:
    \begin{equation*}
        f\left(
            \alpha_0 \prod_{i = 1}^p u_i^{\omega \times X_i} \alpha_i
        \right)
        = P(X_1, \dots, X_p) \quad .
    \end{equation*}
\end{lemma}
\begin{proof}
    Let $\omega$ be an idempotent power for the finite monoid $M$,
    and
    $w \defined \alpha_0 \prod_{i = 1}^p (u_i^{\omega \times X_i} \alpha_i)$.
    Let $k \in \Nat$, and $\pi \colon M^k \topartial \Nat$ be such that
    $f = \pi^\dagger$. That is, 
    $f(w)$ is the sum over all factorizations of $w$
    into $k$ words $(w_1, \dots, w_k) \in \Sigma^*$
    of the value $\pi(\mu(w_1), \dots, \mu(w_k))$.

    Let us define an equivalence relation $\equiv$ over words of
    $\Sigma^*$ as follows: two words $v_1, v_2$ are equivalent if they are
    equal when normalizing them with the rules $u_i^{\omega} u_i^{\omega} \to
    u_i^{\omega}$ for all $1 \leq i \leq p$. That is, we disregard
    repetitions of the factors $u_i^{\omega}$ in the words.
    We lift the equivalence relation $\equiv$ to $k$-tuples of
    words by pointwise application.

    Remark that if $(v_1, \dots, v_k) \equiv (w_1, \dots, w_k)$ then
    $\pi(\mu(v_1), \dots, \mu(v_k)) = \pi(\mu(w_1), \dots, \mu(w_k))$ because
    $\mu(u_i^\omega) = \mu(u_i)^\omega$ is an idempotent element of the monoid
    $M$. In particular, $f(w)$ can be computed as the sum over equivalence
    classes for $\equiv$, of the value of $\pi$ on one representative,
    which belongs to $\Nat$.

    Therefore, it suffices to prove that the number of partitions of $w$ in a
    single equivalence class for $\equiv$ is a polynomial $P(X_1, \dots, X_p)$
    with \kl{positive coefficients} for its \kl{maximal monomials}.

    Notice that to produce all elements in a given equivalence class
    means adding the factors $u_i^\omega$ where $1 \leq i \leq p$
    back to the normalized version
    of the equivalence class. For all elements in an equivalence class,
    one has to add the same amount of $u_i^\omega$ for $1 \leq i \leq p$,
    and therefore the number of possible choices
    is either constant, or 
    is a product of binomials of the form $\binom{X - l}{s}$
    which has the desired shape.
\end{proof}

\begin{corollary}
    \label{n-rat-correct:lem}
    Let $P \in \Rel[X_1, \dots, X_p]$ be \kl{represented}
    by a \kl{$\Nat$-rational series}. Then,
    $P \in \CorrectPoly$.
\end{corollary}
\begin{proof}
    Let $f$ be the \kl{commutative}
    \kl{$\Nat$-rational series}
    with domain $\Sigma \defined \set{a_1, \dots, a_p}$
    that \kl{represents} $P$. Because $f$ has \kl{polynomial growth},
    it is in fact a \kl{$\Nat$-polyregular function}
    (\cref{polynomial-rational-polyreg:fact}).
    Using \cref{n-poly-combinatorics:lem},
    there exists a number $\omega \in \Nat$
    and polynomial $Q$ with \kl{positive coefficients} for its
    \kl{maximal monomials} such that
    for all $X_1, \dots, X_p \geq \omega$:
    \begin{equation*}
        f\left(
            \prod_{i = 1}^p (a_i)^{\omega \times X_i}
        \right)
        = Q(X_1, \dots, X_p)
        = P(\omega X_1, \dots, \omega X_p) 
        \quad .
    \end{equation*}
    We conclude that $P(\omega X_1, \dots, \omega X_p) = Q(X_1, \dots, X_p)$
    as polynomials,
    and in particular that
    the \kl{maximal monomials} of 
    $P$ have \kl{positive coefficients}.

    Remark that for every partial valuation $\nu \colon \vec{X} \topartial \Nat$,
    the polynomial $\restr{P}{\nu}$ continues to be represented
    by a \kl{$\Nat$-polyregular function}, namely
    $f$ partially applied to a word. As a consequence,
    the \kl{maximal monomials} of
    $\restr{P}{\nu}$ also have \kl{positive coefficients}, and 
    we have proven that $P \in \CorrectPoly$.
\end{proof}

\AP The core of the upcoming \cref{lem:correct-to-n-rat} leverages a notion of
\kl{discrete derivative} to perform an induction on the \kl{maximal monomials},
and suitable notions of \kl{translations} of the domain of polynomials. Namely,
we write $\translate{K}$ for the \intro{translation function} that maps a
polynomial $P \in \Rel[X_1, \dots, X_n]$ to the polynomial $P(X_1 + K, \dots,
X_n + K)$.

\begin{fact}
    \label{translation-invariance:fact}
    The following sets are stable under \kl{translations}
    $\translate{K}$,
    where $K \in \Nat$ is non-negative:
    \begin{itemize}
        \item $\CorrectPoly$,
        \item \kl{$\Nat$-rational polynomials},
        \item Polynomials \kl{represented}
            by \kl{star-free $\Nat$-polyregular functions}.
    \end{itemize}
\end{fact}

The idea behind the \kl{discrete derivative} introduced in
\cref{discrete-derivative:def} is to consider the map $P \mapsto P(X) -
P(X-K)$, which will be formulated using the \kl{translation function} in the
case of multiple variables.

\begin{definition}
    \label{discrete-derivative:def}
    Let $\vec{X}$ be a tuple of indeterminates,
    $K \in \Nat$,
    and 
    $P \in \Rel[\vec{X}]$ be a polynomial.
    \begin{equation*}
        \Diff{K}{P} \defined 
        P - \translate{-K}(P) \quad .
    \end{equation*}
\end{definition}

Let us now prove that for a single indeterminate $X$, polynomials $P \in
\CorrectPoly$ are \kl{$\Nat$-rational polynomials}. This will illustrate how
the different ingredients piece together. Given a polynomial $P \in
\CorrectPoly$, the goal is to build a \kl{$\Nat$-polyregular function} $f$ that
\kl{represents} $P$. Note that in this case, the function has unary input and
unary output. We prove the result on the degree $\deg(P)$ of the polynomial
$P$, where the base case is trivial because when $P$ is a constant function,
this constant must be in $\Nat$. For the induction step, we can assume that $P
= \alpha X^n + Q$, where $\deg(Q) < \deg(P)$, $n \geq 1$, and $\alpha > 0$. On
the one hand, $\alpha X^n$ is a \kl{$\Nat$-rational polynomial}, but on the
other hand, $Q$ may not belong to $\CorrectPoly$. However, there exists $K \in
\Nat$ such that $\Diff{K}{X^n} + Q$ belongs to $\CorrectPoly$, for instance by
considering $K$ to be the sum of absolute values of coefficients of $P$.
Furthermore, $\Diff{K}{X^n} + Q$ has degree smaller than $\deg(P)$, hence is a
\kl{$\Nat$-rational polynomial}. Now, $P = \alpha (X - K)^n + \Diff{K}{X^n} +
Q$, and therefore, $P(X+K) = X^n + (\Diff{K}{X^n} + Q)(X+K)$ is a
\kl{$\Nat$-rational polynomial}, \kl{represented} by a \kl{$\Nat$-polyregular
function} $f_1$. Now, for values $k$ of $X$ that are smaller than $K$, we can
compute $P(k) = \restr{P}{X=k}$, which belongs to $\Nat$. It is possible to
build a \kl{$\Nat$-polyregular function} $f_2 \colon \Nat \to \Nat$ that
outputs $P(k)$ if the input is at most $K$, and $f_1(k)$ otherwise.

In the actual statement of \cref{corrected-version:thm}, the representation of
polynomials in $\CorrectPoly$ is strengthened from \kl{$\Nat$-polyregular
functions} to \kl{star-free $\Nat$-polyregular functions}. This is not
surprising, given the conjecture that \kl{ultimate polynomial}
\kl{$\Nat$-polyregular functions} (re-introduced in \cref{star-free:sec}) were
conjectured to be star free in \cite{LOPEZ23b}.

When multiple indeterminate are involved, a bit of care is needed. The key
ingredient is that one has a fine control over the \kl{maximal monomials} of
the \kl{discrete derivatives}, even when fixing some indeterminates.
As one would expect, the \kl{discrete derivatives} are linear operations on
polynomials, that commutes with the partial application operators
$\restr{\cdot}{\nu}$, and the \kl{translation operators} $\translate{K}$.

\begin{fact}
    \label{discrete-deriv-linear:fact}
    For all tuple $\vec{X}$ of indeterminates,
    for all $K \in \Nat$, $L \in \Rel$, for all partial function
    $\nu \colon \vec{X} \topartial \Nat$:
    \begin{enumerate}
        \item $\translate{K} \colon \Rel[\vec{X}] \to \Rel[\vec{X}]$
            is a linear operator,
        \item $\Diff{K}{ \cdot } \colon \Rel[\vec{X}] \to \Rel[\vec{X}]$
            is a linear operator,
        \item $\restr{\cdot}{\nu} \colon \Rel[\vec{X}] \to \Rel[\vec{X}]$
            is a linear operator,
        \item $\Diff{K}{ \cdot } \circ \restr{\cdot}{\nu}
            = \restr{\cdot}{\nu} \circ \Diff{K}{\cdot}$.
        \item $\Diff{K}{ \cdot } \circ \translate{L}
            = \translate{K} \circ \Diff{L}{\cdot}$,
        \item $\restr{\cdot}{\nu} \circ \translate{L}
            = \translate{L} \circ \restr{\cdot}{\translate{L}(\nu)}$.
    \end{enumerate}
\end{fact}

\begin{fact}
    \label{translation-maximal:fact}
    For all $P \in \Rel[\vec{X}]$, and $K \in \Nat$,
    $\MaximalMonomials(P) = \MaximalMonomials(\translate{K}(P))$.
\end{fact}

\AP Let us now introduce the ordering over which the induction of
\cref{lem:correct-to-n-rat} is built. Recall that $\MaximalMonomials(P)$ is the
set of \kl{maximal monomials} of $P$, hence belongs to $\Pfin(\Monomials)$. We
order \kl{monomials} with the \kl{divisibility ordering}, making $\Monomials$ a
\kl{well-quasi-ordering} that is isomorphic to $\Nat^k$ with the product
ordering \cite[see e.g.][Dickson’s Lemma]{SCSC12}. We endow $\Pfin(\Monomials)$
with the \intro{Hoare ordering}, that is, $S_1 \hoareleq S_2$ whenever for all
\kl{monomials} $M_1 \in S_1$, there exists a monomial $M_2 \in S_2$, such that
$M_1$ \kl{divides} $M_2$. The set $(\Pfin(\Monomials), \hoareleq)$ remains a
\kl{well-quasi-ordering} \cite[see e.g.][Hoare quasi-ordering]{SCSC12}.

The following \cref{derivation-simplifies:lemma,derivation-covers:lemma} 
show how the \kl{discrete differentiation} operator
interacts with \kl{maximal monomials} with respect to the \kl{Hoare ordering}:
it extracts \emph{submaximal} monomials from a given polynomial.

\begin{lemma}
    \label{derivation-simplifies:lemma}
    For all $P \in \Rel[\vec{X}]$ that are non-constant,
    for all $K \in \Nat$,
    $\MaximalMonomials(\Diff{K}{P}) \hoarele
    \MaximalMonomials(P)$.
\end{lemma}
\begin{proof}
    By linearity of $\Diff{K}{P}$ (\cref{discrete-deriv-linear:fact}), 
    it suffices to prove the result
    for \kl{monomials} $S \in \Monomials$.
\end{proof}

\begin{fact}
    Let $K > 0$,
    and let $M,T$ be two \kl{monomials}, such that
    $T$ \kl{strictly divides} $M$.
    Then,
    $T$ \kl{divides} some \kl{maximal monomial}
    of $\Diff{K}{M}$.
\end{fact}

\begin{lemma}
    \label{derivation-covers:lemma}
    For all $P,Q \in \Rel[\vec{X}]$ that are non-constant,
    for all $K > 0$,
    if $\MaximalMonomials(P) \hoarele \MaximalMonomials(Q)$,
    then
    $\MaximalMonomials(P) \hoareleq \MaximalMonomials(\Diff{K}{Q})$.
\end{lemma}

\begin{lemma}
    \label{derivation-stabilises-correct:lem}
    Let $P \in \CorrectPoly$,
    $P_1$ be the sum of \kl{maximal monomials} of $P$,
    and $P_2 \defined P - P_1$ be the sum of
    non-maximal monomials of $P$.
    There exists a computable $K$,
    such that
    $(\Diff{K}{P_1} + P_2) \in \CorrectPoly$.
\end{lemma}

@@

\begin{lemma}
    \label{decidability-correct:lem}
    It is decidable whether a polynomial 
    $P \in \Rel[\vec{X}]$
    belongs to $\CorrectPoly$.
\end{lemma}
\begin{proof}
    Let $P \in \Rel[\vec{X}]$. We prove by induction on
    $\MaximalMonomials(P)$ that whether $P \in \CorrectPoly$
    is decidable.

    If $P$ is a constant polynomial, then
    $P \in \CorrectPoly$ if and only if $P = n \in \Nat$
    which is decidable.

    Otherwise, $P$ is not a constant polynomial, and let us write $P = P_1 +
    P_2$ where $P_1$ is the sum of the \kl{maximal monomials} of $P$. If $P_1
    \not \in \Nat[\vec{X}]$, then $P \not \in \CorrectPoly$, and this is
    decidable. Hence, we can assume that $P_1 \in \Nat[\vec{X}]$. Using
    \cref{derivation-stabilises-correct:lem}, there exists a computable bound
    $K$, such that if $P \in \CorrectPoly$, then t $Q \defined (\Diff{K}{P_1} +
    P_2) \in \CorrectPoly$. Leveraging, \cref{derivation-simplifies:lemma}, we
    also know that $\MaximalMonomials(Q) \hoarele \MaximalMonomials(P)$, and
    therefore that we can decide whether $Q \in \CorrectPoly$. Notice that if
    $Q \not \in \CorrectPoly$, then $P \not \in \CorrectPoly$, hence we can
    assume that $Q \in \CorrectPoly$.

    Similarly, for partial valuations
    $\nu \colon \vec{X} \topartial \set{0, \dots, K}$
    fixing at least one indeterminate, $\MaximalMonomials(\restr{P}{\nu})
    \hoarele \MaximalMonomials{P}$. 
    We can check for all these partial valuations that
    $\restr{P}{\nu} \in \CorrectPoly$ by induction hypothesis,
    and if one test is not satisfied, then $P \not \in \CorrectPoly$.

    Let us now conclude that if all the above checks answered correctly,
    then $P \in \CorrectPoly$. Let $\nu \colon \vec{X} \topartial \Nat$
    be a partial valuation. Then, 
    one of the three following cases arise:
    \begin{itemize}
        \item 
            Either,
        $\nu$ fixes \emph{some} indeterminate $X_i$ to a value $n_i < K$.
        In that case, let us define $\delta$ as the partial function that coincides
        with $\nu$ on its domain
        $\dom(\delta) = \dom(\nu) \setminus \set{ X_i }$.
        Then, $\restr{P}{\nu} = \restr{\restr{P}{X_i = n_i}}{\delta}$.
        As we previously checked that $\restr{P}{X_i = n_i} \in \CorrectPoly$,
        we conclude in particular that 
        $\restr{P}{\nu}$ has \kl{non-negative} \kl{maximal monomials}.

        \item Or, for all $X_i$ in the (potentially empty) domain of $\nu$, $\nu(X_i)
            \geq K$. In that case, let us write $\delta \defined
            \translate{-K}(\nu)$, that is the partial map $\delta \colon
            \vec{X} \topartial \Nat$ having the same domain as $\nu$, and such
            that for all $X_i \in \dom(\nu)$, $\delta(X_i) \defined \nu(X_i) -
            K$. Recall that $Q = \Diff{K}{P_1} + P_2$, and therefore that
            $\translate{K}(P) = P_1 + \translate{K}(Q)$. By linearity,
            $\restr{\translate{K}(P)}{\delta} = \restr{P_1}{\delta} +
            \restr{\translate{K}(Q)}{\delta}$, and we conclude that the
            \kl{maximal monomials} of $\restr{\translate{K}(P)}{\delta}$ are
            \kl{non-negative}.
            Since $\restr{\translate{K}(P)}{\delta} = \translate{K}{\restr{P}{\nu}}$,
            and since
            $\translate{K}$ leaves \kl{maximal monomials} invariant,
            we conclude that
            the \kl{maximal monomials} of $\restr{P}{\nu}$
            are the same as those of $\restr{\translate{K}(P)}{\delta}$,
            and are therefore \kl{non-negative}.
            \qedhere
    \end{itemize}
\end{proof}

@@

\begin{lemma}
    \label{lem:correct-covered-2}
    $\CorrectPoly[X,Y] = \CoveredPoly[X,Y]$.
\end{lemma}
\begin{proof}
    It is clear that $\CorrectPoly[X,Y] \subseteq \CoveredPoly[X,Y]$,
    by considering the empty valuation $\nu \colon \set{X,Y} \topartial \Nat$.
    For the converse inclusion, let us consider $P(X,Y)$
    that is \kl{non-negative}, such that the \kl{maximal monomials}
    have \kl{positive coefficients}.

    If we fix none of the variables, then the \kl{maximal monomials}
    have \kl{positive coefficients} by assumption. If we fix one of the
    variables, we can assume without loss of generality that we 
    fix $X = k$ for some $k \in \Nat$.
    Then $P(k,Y)$ is a \kl{non-negative} \emph{univariate} polynomial, 
    and therefore must have a positive leading coefficient
    (which is the unique \kl{maximal monomial} in this case)
    or be constant equal to 0. In both cases, the \kl{maximal monomials}
    have \kl{positive coefficients}.
    The same reasoning applies \emph{a fortiori} in the case where
    we fix the two indeterminate, leading to a constant polynomial.
\end{proof}

It was conjectured in the 
\cite[Open question 5.55]{gaetanphd}.
\cite[Conjecture 7.61]{gaetanphd}
Let us remark that a previous conjecture was not true, and provide the
recursively defined variant. This can be understood in the light of
\cref{corrected-version:thm}, as the fact that \kl{$\Nat$-rational polynomials}
are essentially $\Nat$-polynomials.

\begin{example}
    The polynomial $\BadPoly$ is not a 
    \kl{$\Nat$-rational polynomial},
    but is \kl{non-negative} and satisfies
    $\translate{10}(\BadPoly) \in \Nat[\vec{X}]$.
\end{example}

\begin{lemma}
    \label{derivation-translation:lem}
    Let $P \in \Rel[\vec{X}]$, the following are equivalent
    \begin{enumerate}
        \item \label{d-t-correct:item} $P \in \CorrectPoly$,
        \item \label{d-t-transl:item}
            There exists $K \in \Nat$,
            such that for every partial
            function $\nu \colon \vec{X} \topartial \Nat$,
            $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    To prove \cref{d-t-transl:item} $\implies$
    \cref{d-t-correct:item}, let us remark that
    if $P$ is \kl{non-negative} and $\translate{K}(P) \in \Nat[\vec{X}]$,
    then $\translate{K}(P) \in \CorrectPoly$ thanks to
    \cref{n-poly-n-poly:example}.
    In particular, 
    then for every partial valuation 
    $\nu \colon \vec{X} \topartial \Nat$, 
    $\restr{\translate{K}(P)}{\nu} \in \Nat[\vec{X}]$,
    and in particular conclude that
    \kl{maximal monomials} of $\restr{\translate{K}(P)}{\nu}$ are exactly the
    \kl{maximal monomials} of $\restr{P}{\nu}$, the former being
    \kl{non-negative} if $\translate{K}(P) \in \Nat[\vec{X}]$.

    Conversely, the implication \cref{d-t-correct:item} $\implies$
    \cref{d-t-transl:item} is proven by induction on $\MaximalMonomials(P)$.
    If $P$ is a constant polynomial, then $P = n$ for some $n \in \Nat$,
    and we conclude that $\translate{0}(P) \in \Nat[\vec{X}]$.
    Otherwise, $P = P_1 + P_2$ where $P_1$ is the sum of the \kl{maximal
    monomials} of $P$.
    Using \cref{derivation-stabilises-correct:lem},
    there exists a computable $K \in \Nat$
    such that 
    $Q \defined
    \Diff{K}{P_1} + P_2$ belongs to $\CorrectPoly$.
    Since, $\MaximalMonomials(Q) \hoarele \MaximalMonomials(P)$,
    we conclude by induction hypothesis that
    there exists $L \in \Nat$ such that 
    for every partial valuation $\nu \colon \vec{X} \topartial \Nat$,
    $\translate{L}(\restr{Q}{\nu}) \in \Nat[\vec{X}]$. 
    Similarly,
    for all partial functions $\nu \colon \vec{X} \topartial \Nat$
    fixing at least one indeterminate (appearing in $P$),
    $\MaximalMonomials(\restr{P}{\nu}) \hoarele \MaximalMonomials(P)$,
    hence there exists $L_\nu \in \Nat$ such that
    $\translate{L_\nu}{\restr{P}{\nu}} \in \Nat[\vec{X}]$.
    Let us define $L_m$ to be the maximum of $L$, and $L_\nu$
    where $\nu$ ranges over partial functions described above.

>>>>

The counter example provided by \cref{def:bad-polynomial} relies on the fact
that $\CoveredPoly$ is not stable under fixing indeterminates, while
\kl{$\Nat$-polyregular functions} are. In this section, we prove that closing
$\CorrectPoly$ under assignments of variable is enough to recover from
\cref{karh:thm}.

@@

Let us now prove that \kl{$\Nat$-rational polynomials} are in $\CorrectPoly$.
This follows from the correct implication in the statement of \cref{karh:thm},
but we provide a self-contained proof using a refinement of the
classical combinatorial arguments for $\ZPoly$ \cite[Lemma 4.16]{LOPEZ23b} and
$\NPoly$ \cite[Lemma 5.37]{gaetanphd}.

\begin{lemma}
    \label{n-poly-combinatorics:lem}
    Let $f$ be a  \kl{$\Nat$-polyregular} function. 
    There exists $\omega \in \Nat$
    such that for all $p \in \Nat$,
    for all $\alpha_0, \dots, \alpha_p \in \Sigma^*$,
    for all $u_1, \dots, u_p \in \Sigma^*$,
    there exists a polynomial $P \in \Rel[X_1, \dots, X_p]$
    whose \kl{maximal monomials} are \kl{non-negative},
    and such that for all $X_1, \dots, X_p \geq \omega$:
    \begin{equation*}
        f\left(
            \alpha_0 \prod_{i = 1}^p u_i^{\omega X_i} \alpha_i
        \right)
        = P(X_1, \dots, X_p) \quad .
    \end{equation*}
\end{lemma}

\begin{corollary}
    \label{n-rat-correct:lem}
    Let $P \in \Rel[X_1, \dots, X_p]$ be \kl{represented}
    by a \kl{$\Nat$-rational series}. Then,
    $P \in \CorrectPoly$.
\end{corollary}
\begin{proof}
    Let $f$ be the \kl{commutative}
    \kl{$\Nat$-rational series}
    with domain $\Sigma \defined \set{a_1, \dots, a_p}$
    that \kl{represents} $P$. Because $f$ has \kl{polynomial growth},
    it is in fact a \kl{$\Nat$-polyregular function}
    (\cref{polynomial-rational-polyreg:fact}).
    Using \cref{n-poly-combinatorics:lem},
    there exists a number $\omega \in \Nat$
    and polynomial $Q$ with \kl{non-negative}
    \kl{maximal monomials} such that
    for all $X_1, \dots, X_p \geq \omega$:
    \begin{equation*}
        f\left(
            \prod_{i = 1}^p a_i^{\omega X_i}
        \right)
        = Q(X_1, \dots, X_p)
        = P(\omega X_1, \dots, \omega X_p) 
        \quad .
    \end{equation*}
    We conclude that $P(\omega X_1, \dots, \omega X_p) = Q(X_1, \dots, X_p)$
    as polynomials,
    and in particular that
    the \kl{maximal monomials} of 
    $P$ are \kl{non-negative}.

    Remark that for every partial valuation $\nu \colon \vec{X} \topartial \Nat$,
    the polynomial $\restr{P}{\nu}$ continues to be represented
    by a \kl{$\Nat$-polyregular function}, namely
    $f$ partially applied to a word. As a consequence,
    the \kl{maximal monomials} of
    $\restr{P}{\nu}$ are also \kl{non-negative}, 
    and
    we have proven that $P \in \CorrectPoly$.
\end{proof}

\AP The core of the upcoming \cref{lem:correct-to-n-rat} leverages a notion of
\kl{discrete derivative} to perform an induction on the \kl{maximal monomials}.
This notion of derivation is built by translating the domain of the polynomial.
To that end, let us write $\translate{K}$ for the \intro{translation function}
that maps a polynomial $P \in \Rel[X_1, \dots, X_n]$ to the polynomial $P(X_1 +
K, \dots, X_n + K)$.

\begin{definition}
    \label{discrete-derivative:def}
    Let $\vec{X}$ be a tuple of indeterminates,
    $K \in \Nat$,
    and 
    $P \in \Rel[\vec{X}]$ be a polynomial.
    \begin{equation*}
        \Diff{K}{P} \defined 
        \translate{K}(P) - P
        \quad .
    \end{equation*}
\end{definition}

As one would expect, the \kl{discrete derivatives} are linear operations on
polynomials, that commutes with \kl{translation operators}. However, the
\kl{translations operators} do not perfectly commute with partial applications
$\restr{\cdot}{\nu}$.

\begin{fact}
    \label{discrete-deriv-linear:fact}
    For all tuples $\vec{X}$ of indeterminates,
    for all $K \in \Nat$, $L \in \Rel$, for all partial function
    $\nu \colon \vec{X} \topartial \Nat$:
    \begin{enumerate}
        \item The maps $\translate{K} \colon \Rel[\vec{X}] \to \Rel[\vec{X}]$,
        $\Diff{K}{ \cdot } \colon \Rel[\vec{X}] \to \Rel[\vec{X}]$,
        and
        $\restr{\cdot}{\nu} \colon \Rel[\vec{X}] \to \Rel[\vec{X}]$
            are linear operators,
        \item $\Diff{K}{ \cdot } \circ \translate{L}
            = \translate{K} \circ \Diff{L}{\cdot}$,
        \item $\restr{\cdot}{\nu} \circ \translate{L}
            = \translate{L} \circ \restr{\cdot}{\translate{L}(\nu)}$.
    \end{enumerate}
\end{fact}

To build some intuition on the proof to come, let us sketch the argument in the
case of a single indeterminate $X$. Given a polynomial $P \in \CorrectPoly$,
the goal is to build a \kl{$\Nat$-polyregular function} $f$ that
\kl{represents} $P$. Note that in this case, the function has unary input and
unary output. We prove the result by induction on the degree $\deg(P)$ of the
polynomial $P$, where the base case is trivial because when $P$ is a constant
function, this constant must be in $\Nat$. For the induction step, we can
assume that $P = \alpha X^n + Q$, where $\deg(Q) < \deg(P)$, $n \geq 1$, and
$\alpha > 0$. On the one hand, $\alpha X^n$ is a \kl{$\Nat$-rational
polynomial}, but on the other hand, $Q$ may not belong to $\CorrectPoly$.
However, there exists $K \in \Nat$ such that $\Diff{K}{X^n} + \translate{K}(Q)$
belongs to $\CorrectPoly$, for instance by considering $K$ to be the sum of
absolute values of coefficients of $P$. Furthermore, $\Diff{K}{X^n} +
\translate{K}(Q)$ has degree smaller than $\deg(P)$, hence is a
\kl{$\Nat$-rational polynomial}. Now, $\translate{K}(P) = \alpha X^n +
\Diff{K}{X^n} + \translate{K}(Q)$ is a \kl{$\Nat$-rational polynomial},
represented by a function $f_1$. Now, for values $k$ of $X$ that are smaller
than $K$, we can compute $P(k) = \restr{P}{X=k}$, which belongs to $\Nat$. It
is possible to build a \kl{$\Nat$-polyregular function} $f_2 \colon \Nat \to
\Nat$ that outputs $P(k)$ if the input is at most $K$, and $f_1(k - K)$
otherwise.

In the actual statement of \cref{corrected-version:thm}, the representation of
polynomials in $\CorrectPoly$ is strengthened from \kl{$\Nat$-polyregular
functions} to \kl{star-free $\Nat$-polyregular functions}. This is not
surprising, given the conjecture that \kl{ultimately polynomial}
\kl{$\Nat$-polyregular functions} (re-introduced in \cref{star-free:sec}) were
conjectured to be star free (\cref{zsf-nsf:conjecture}), and that polynomials
are always \kl{ultimately polynomial}.

When multiple indeterminate are involved, the notion of 
degree must be adapted, and to ensure that the decision procedure
is effective, one must obtain explicit bounds on the
translations that are needed.

\AP To that end, let us now introduce the ordering over which the induction of
\cref{lem:correct-to-n-rat} is built. Recall that $\MaximalMonomials(P)$ is the
set of \kl{maximal monomials} of $P$, hence belongs to $\Pfin(\Monomials)$. We
order \kl{monomials} with the \kl{divisibility ordering}, making $\Monomials$ a
\kl{well-quasi-ordering} that is isomorphic to $\Nat^k$ with the product
ordering \cite[see e.g.][Dickson’s Lemma]{SCSC12}. We endow $\Pfin(\Monomials)$
with the \intro{Hoare ordering}, that is, $S_1 \hoareleq S_2$ whenever for all
\kl{monomials} $M_1 \in S_1$, there exists a monomial $M_2 \in S_2$, such that
$M_1$ \kl{divides} $M_2$. The set $(\Pfin(\Monomials), \hoareleq)$ remains a
\kl{well-quasi-ordering} \cite[see e.g.][Hoare quasi-ordering]{SCSC12}.

\begin{fact}
    \label{maxi-monomials-submodular:fact}
    Let $P, Q$ be two polynomials.
    Then, $\MaximalMonomials(P+Q) \hoareleq \MaximalMonomials(P)
    \cup \MaximalMonomials(Q)$.
\end{fact}

Let us now illustrate how the \kl{discrete differentiation} operator interacts
with \kl{maximal monomials} with respect to the \kl{Hoare ordering}: it
extracts \emph{submaximal} monomials from a given polynomial.

\begin{fact}
    \label{translation-maximal:fact}
    For all $P \in \Rel[\vec{X}]$, and $K \in \Nat$,
    \begin{equation*}
        \MaximalMonomials(P) = \MaximalMonomials(\translate{K}(P))
        \quad .
    \end{equation*}
\end{fact}

\begin{fact}
    \label{derivation-monomials:fact}
    Let $K > 0$,
    and let $M,T$ be two \kl{monomials}, such that
    $T$ \kl{strictly divides} $M$.
    Then,
    every \kl{monomial} of $\Diff{K}{M}$ \kl{strictly divides} $M$,
    and 
    $T$ \kl{divides} some \kl{maximal monomial}
    of $\Diff{K}{M}$, which has a coefficient
    that is a multiple of $K$.
\end{fact}

\begin{lemma}
    \label{derivation-simplifies:lemma}
    For all $P \in \Rel[\vec{X}]$ that are non-constant,
    for all $K \in \Nat$,
    $\MaximalMonomials(\Diff{K}{P}) \hoarele
    \MaximalMonomials(P)$.
\end{lemma}

The following \cref{derivation-stabilises-correct:lem} is the main
combinatorial argument of this section. Note that we must be careful
about it. \textbf{TODO: write a nice thing here!}.

\begin{lemma}
    \label{derivation-stabilises-correct:lem}
    Let $P \in \CorrectPoly$,
    $P_1$ be the sum of \kl{maximal monomials} of $P$,
    and $P_2 \defined P - P_1$ be the sum of
    non-maximal monomials of $P$.
    There exists a computable $K$,
    such that
    $Q \defined (\Diff{K}{P_1} + \translate{K}(P_2)) \in \CorrectPoly$.
\end{lemma}

@@

\begin{lemma}
    \label{decidability-correct:lem}
    It is decidable whether a polynomial 
    $P \in \Rel[\vec{X}]$
    belongs to $\CorrectPoly$.
\end{lemma}

@@

\begin{lemma}
    \label{lem:correct-covered-2}
    $\CorrectPoly[X,Y] = \CoveredPoly[X,Y]$.
\end{lemma}
\begin{proof}
    It is clear that $\CorrectPoly[X,Y] \subseteq \CoveredPoly[X,Y]$,
    by considering the empty valuation $\nu \colon \set{X,Y} \topartial \Nat$.
    For the converse inclusion, let us consider $P(X,Y)$
    that is \kl{non-negative}, such that the \kl{maximal monomials}
    are  \kl{non-negative}.

    If we fix none of the variables, then the \kl{maximal monomials}
    are \kl{non-negative} by assumption. If we fix one of the
    variables, we can assume without loss of generality that we 
    fix $X = k$ for some $k \in \Nat$.
    Then $P(k,Y)$ is a \kl{non-negative} \emph{univariate} polynomial, 
    and therefore must have a positive leading coefficient
    (which is the unique \kl{maximal monomial} in this case)
    or be constant equal to 0. In both cases, the \kl{maximal monomials}
    have positive coefficient.
    The same reasoning applies \emph{a fortiori} in the case where
    we fix the two indeterminate, leading to a constant polynomial.
\end{proof}

Let us also state that polynomials in $\CorrectPoly$ are, up to suitable
translations, polynomials in $\Nat[\vec{X}]$. Beware that partial applications
are still needed in this characterization, as \cref{bad-poly-translate:ex}
illustrates.

\begin{lemma}
    \label{derivation-translation:lem}
    Let $P \in \Rel[\vec{X}]$, the following are equivalent:
    \begin{enumerate}
        \item \label{d-t-correct:item} $P \in \CorrectPoly$,
        \item \label{d-t-transl:item}
            There exists $K \in \Nat$,
            such that for every partial
            function $\nu \colon \vec{X} \topartial \Nat$,
            $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$.
    \end{enumerate}
\end{lemma}

\begin{example}
    \label{bad-poly-translate:ex}
    The polynomial $\BadPoly$ is not a 
    \kl{$\Nat$-rational polynomial},
    but is \kl{non-negative} and satisfies
    $\translate{10}(\BadPoly) \in \Nat[\vec{X}]$.
\end{example}
